Tutorial: SQL-on-Hadoop Systems

### 2015-04-02 paper.tex
Some edits on Impala, and updated refs

@@

Cloudera Impala \cite{Kornacker+15} an open-source, fully-integrated MPP SQL query engine. Impala's goal is to combine the familiar SQL support and multi-user performance of a traditional analytic database with the scalability and flexibility of Apache Hadoop. Unlike other systems (often forks of Postgres), Impala is a brand-new engine. It is written from the ground up in C++ and Java, and employs uses LLVM to generate code at runtime in order to speed up frequently executed code paths. It maintains Hadoop's flexibility by utilizing standard components (HDFS, HBase, Metastore, YARN, Sentry) and is able to read the majority of the widely-used file formats (e.g. Parquet, Avro, RCFile). To reduce latency, such as that incurred from utilizing MapReduce or by reading data remotely, Impala implements a distributed architecture based on daemon processes that are responsible for all aspects of query execution and that run on the same machines as the rest of the Hadoop infrastructure. The result is performance that is on par or exceeds that of commercial MPP analytic DBMSs, depending on the particular workload.

>>>>

Most file formats are variants of PAX \cite{AilamakiDHS01}

@@

Cloudera Impala \cite{Kornacker+15} is an open-source, fully-integrated MPP SQL query engine. Unlike other systems (often forks of Postgres), Impala is a brand-new engine. It is written from the ground up in C++ and Java. To reduce latency, such as that incurred from utilizing MapReduce or by reading data remotely, Impala implements a distributed architecture based on daemon processes that are responsible for all aspects of query execution and that run on the same machines as the rest of the Hadoop infrastructure. 
In order to perform data scans from both disk and memory at or near hardware speed,
Impala uses an HDFS feature called {\em short-circuit local reads} \cite{hdfs:shortcircuit} to bypass the DataNode protocol when reading from local disk. Impala can read at almost disk bandwidth and is typically able to saturate all available disks. Furthermore, {\em HDFS caching} \cite{hdfs:caching} allows Impala to access memory-resident data at memory bus speed and also saves CPU cycles as there is no need to copy data blocks and/or checksum them.
Impala leverages decades of research in parallel databases. The execution model is the traditional
Volcano-style with Exchange operators \cite{Graefe90}. Processing is performed batch-at-a-time:
each \texttt{GetNext()} call operates over batches of rows, similar to
\cite{PadmanabhanMAJ01}. A main characteristic of Impala is that employs LLVM to generate code at runtime to speed up frequently executed code paths \cite{WandermanL14}. The result is performance that is on par or exceeds that of commercial MPP analytic DBMSs, depending on the particular workload.
